### 2015-11-07 Assignment IV/Assignment4-Ditz_Schroeder-WS15-16.tex
add die bib datei für unsere referenzen und neueste version vom tex file

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
\usepackage{color}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{grffile} %allows to use pngs
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
%\usepackage{TikZ} %phylogenetischer Baum
%\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
%\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

@@

\section*{Theoretical Assignment - \textsl{Progressive alignment}}
Progressive alignment methodes are one way to create multiple sequence alignments (MSA), in this task we had to accomplish manualy a progressive alignment. The first step for the progressive alignment was to generate the pairwise global distance matrices. Afterwards the global optima were entered into a new table, which was used to generate the guide tree in the second step.(figure \ref{DP2}).
\begin{figure}[h]
	\includegraphics[width=\textwidth]{Img/Exercise2-DP-Matrix.png}
	\caption{ Global alignments lead to the basis table used to generate the guide tree}
	\label{DP2}
\end{figure}

@@

\begin{figure}[h]
	\includegraphics[width=\textwidth]{Img/Exercise2-Guidetree.png}
	\caption{ The guide tree is made from the final table from step 1}
	\label{guidetree2}
\end{figure}

The last step was creating the actual MSA. The 'complete alignment' methode was used. This means after every step, the distance of the new sequences was determined to each other sequence in the cluster. The optimal resulting score was used to get the best alignment for the new Sequence. after 3 steps the MSA was completed. (figure \ref{completealign})
\begin{figure}[h]
	\includegraphics[width=\textwidth]{Img/Exercise2-CompleteAlignment.png}
	\caption{ Complete alignment methode is used to get an MSA from the guidetree from step 2	}
	\label{completealign}
\end{figure}
\section*{Practical Assignment - \textsl{Comparing multiple alignment}}

>>>>

\usepackage[T1]{fontenc} % T1 Schrift Encoding
\usepackage{textcomp}    % Zusatzliche Symbole (Text Companion font extension)
\usepackage{lmodern,dsfont}     % Latin Modern Schrift
\usepackage{dsfont}
\usepackage{color}
%\usepackage{wasysym}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{grffile} %allows to use pngs
\usepackage{eurosym}
%\usepackage{txfonts}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{listings}
\usepackage{etextools}
\usepackage{ifthen}
\usepackage{cite}
%\usepackage{TikZ} %phylogenetischer Baum
%\usetikzlibrary{calc, shapes, backgrounds} %für die Phylogenetische bäume
%\usetikzlibrary{automata,arrows}
\usepackage{subfigure}

@@

\section*{Theoretical Assignment - \textsl{Progressive alignment}}
Progressive alignment methodes are one way to create multiple sequence alignments (MSA), in this task we had to accomplish manualy a progressive alignment. The first step for the progressive alignment was to generate the pairwise global distance matrices. Afterwards the global optima were entered into a new table, which was used to generate the guide tree in the second step.(figure \ref{DP2}).
\begin{figure}[ht]
        \centering
	\includegraphics[width=0.7\textwidth]{Img/Exercise2-DP-Matrix.png}
	\caption{ Global alignments lead to the basis table used to generate the guide tree}
	\label{DP2}
\end{figure}

@@

\begin{figure}[ht]
        \centering
	\includegraphics[width=0.7\textwidth]{Img/Exercise2-Guidetree.png}
	\caption{ The guide tree is made from the final table from step 1}
	\label{guidetree2}
\end{figure}

The last step was creating the actual MSA. The 'complete alignment' methode was used. This means after every step, the distance of the new sequences was determined to each other sequence in the cluster. The optimal resulting score was used to get the best alignment for the new Sequence. after 3 steps the MSA was completed. (figure \ref{completealign})
\begin{figure}[ht]
        \centering
	\includegraphics[width=0.7\textwidth]{Img/Exercise2-CompleteAlignment.png}
	\caption{ Complete alignment methode is used to get an MSA from the guidetree from step 2	}
	\label{completealign}
\end{figure}

\section*{Practical Assignment - \textsl{Comparing multiple alignment}}
Multiple sequence alignments (MSA) are very important to detect similarities between several 
sequences and e.g. determine whether a sequence belongs to a certain family or not. Hence, many 
different programs, such as ClustalW, were developed to calculate a MSA. We calculated a multiple 
sequence alignment for three files (BB11007, BB20002 and Prolyl-tRNA), each file contained several 
sequences, with different programs 
to compare the results. The used programs are MAFFT\cite{Katoh1}, MUSCLE\cite{Edgar1, Edgar2} and 
Clustal$\Omega$\cite{Sievers539}. The resulting MSAs were compared with a gold standard provided by 
BaliBase\cite{Bahr1}. This database provide a large number of reference MSAs as well as a program 
\grqq BaliScore\grqq , which can be used to calculate different scores to evaluate your MSA. For 
this assignment we are interested in the total column score, which is the percentage of accurately 
reconstructed columns of the reference MSA, and the sum-of-pair score, which is for BaliScore the 
percentage of pairs of aligned residues that are similar in the reference and the reconstructed 
MSA. Furthermore, a important information is the runtime of different programs. We used the bash 
command \textit{time} to collect this information.\\
\\
\noindent The resulting total column scores can be found in figure \ref{totCol}. All calculated sum-
of-pairs scores are written down in figure \ref{SPscore} and the runtime of all programs can be found 
in figure \ref{runtime}.

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
  \textit{total column score} & \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & $c_{1,1}$ & $c_{2,1}$ & $c_{3,1}$ \\
  \textbf{BB20002} & $c_{1,2}$ & $c_{2,2}$ & $c_{3,2}$ \\
  \textbf{Prolyl-tRNA} & $c_{1,3}$ & $c_{2,3}$ & $c_{3,3}$ \\
 \end{tabular}}
 \caption{total column score of all used programs for example files}
 \label{totCol}
\end{figure}

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
   \textit{SP-score}& \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & $s_{1,1}$ & $s_{2,1}$ & $s_{3,1}$ \\
  \textbf{BB20002} & $s_{1,2}$ & $s_{2,2}$ & $s_{3,2}$ \\
  \textbf{Prolyl-tRNA} & $s_{1,3}$ & $s_{2,3}$ & $s_{3,3}$ \\
 \end{tabular}}
 \caption{sum-of-pairs score of all used programs for example files}
 \label{SPscore}
\end{figure}

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
  \textit{runtime} & \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & $t_{1,1}$ & $t_{2,1}$ & $t_{3,1}$ \\
  \textbf{BB20002} & $t_{1,2}$ & $t_{2,2}$ & $t_{3,2}$ \\
  \textbf{Prolyl-tRNA} & $t_{1,3}$ & $t_{2,3}$ & $t_{3,3}$ \\
 \end{tabular}}
 \caption{runtime of all used programs for example files}
 \label{runtime}
\end{figure}

\newpage
\bibliography{Assignment4-Ditz_Schroeder-WS15-16.bib}
\bibliographystyle{ieeetr}
